Results publications are inadequately linked to trial registrations: An automated pipeline and evaluation of German university medical centers

### 2021-07-22 docs/manuscript.Rmd
Add logistic regression to manuscript

# Source analyses
source(here("R", "functions", "screening.R"))
# # source(here("R", "functions", "report-results.R"))
source(here("R", "analysis", "01_screen-trials.R"))
# # source(here("R", "analysis", "02_flowchart-trials.R"))
source(here("R", "analysis", "03_tbl-trials.R"))
source(here("R", "analysis", "04_link-reg-pub.R"))
# # source(here("R", "analysis", "05_upset-link-reg-pub.R"))
source(here("R", "analysis", "06_plot-change-over-time.R"))
source(here("R", "analysis", "trn-regex-analysis.R"))
source(here("R", "analysis", "cross-registrations.R"))
# # fs::dir_ls(here("R", "analysis")) %>% 
# #   purrr::walk(source)
```

@@

**Secondary analyses** We explored<!--# tested --> the relationships between (1) TRN in abstract or full-text, and TRN in PubMed metadata, as well as between (2) TRN in PubMed metadata and a linked publication in the registration. We also explored the relationship between registration-publication links and (1) registry, and (2) prospective registration <!--# not sure about this. i think if we do this then glm for the various practices -->. We looked at change in registration-publication linking practices over time. We also looked at reported cross-registrations in the registries as well as additional TRNs reported in trial publications.<!--# We also looked at reporting rates by UMCs, randomized controlled trials (RCTs), and ICMJE-status. We also explored the association of TRN reporting practices with other responsible trial research practices, namely prospective registration and summary results reporting. All analyses are exploratory. -->

@@

## Automated pipeline for data sources and collection

@@

### Exclusion/inclusion criteria

@@

<!--# Using registry APIs, when available, and web scraping, we checked whether each unique TRN resolved to a valid trial registration. currently placeholder "TRUE" such that all TRNs resolve in the registries --><!--# Since i'm focusing iv trn only, should i change this section so not ALL the trns? then would have to change my backend work. or can keep this section but maybe not report all the details in text but just appendix? -->

@@

The IntoValue trials dataset included the full sample of trials led by a German UMC that were registered on ClinicalTrials.gov or DRKS and completed between 2009 and 2017 (n = `r n_trials_iv`). After applying our exclusion criteria, our sample included `r n_trials_ft_pdf` trials with `r n_pubs_unique` unique results publications indexed in PubMed and available as full-text, as some publications include results from more than one trial. Figure \@ref(fig:flow-trials) details the trial and publication screening. Table \@ref(tab:tbl-trials) shows summary descriptive details of the trials with results publications. <!--# QUESTION: what summary info about trials and publications? I also include a long version of demographic data in case we want to include any of these, ideally with collapsed categories, in Table @ref(tab:tbl-trials-full) -->

@@

While per our inclusion criteria all trials in our sample (n = `r n_trials_ft_pdf`) were registered and had a publication, only `r report_n_pct_trials(n_trials_link_full)` trials had the most comprehensive registration-publication linking, meaning a linked publication in the registration as well as the TRN in the publication full-text, abstract, and metadata. In contrast, we found `r report_n_pct_trials(n_trials_link_none)` trials with no links in either the registration or the publication. Figure \@ref(fig:plot-upset-links-reg-pub-registries) shows the breakdown of registration-publication links across the various components for ClinicalTrial.gov and DRKS trials.

@@

@@

We found that `r report_pct_num_denom_trials(n_trials_link_reg)` of registrations included a reference with a publication identifier (DOI or PMID) to the associated publication in our sample of trials with published results. Only `r report_pct_num_denom_trials(n_trials_link_reg_drks, n_trials_drks)` of DRKS trials referenced the results publication in the registration, in contrast with `r report_pct_num_denom_trials(n_trials_link_reg_ctgov, n_trials_ctgov)` of ClinicalTrials.gov trials, of which `r report_pct_num_denom_trials(n_trials_ctgov_link_reg_trn_si_derived, n_trials_link_reg_ctgov)` were automatically indexed.

@@

We found that trials with a ClinicalTrials.gov TRN were significantly more likely than trials with a DRKS TRN to include a publication reference in the registry (`r apa::apa(chisq_link_by_reg_reg, print_n = TRUE)`) and to include the TRN in the PubMed metadata (`r apa::apa(chisq_link_by_reg_si, print_n = TRUE)`) and abstract (`r apa::apa(chisq_link_by_reg_abs, print_n = TRUE)`). We did not find a significant relationship between registry and TRN reporting in full-text (`r apa::apa(chisq_link_by_reg_pdf, print_n = TRUE)`). Table \@ref(tab:tbl-link-by-reg) shows summary registration-publication links by registry.

<!--# point estimtes in table should have CI, right? -->

@@

### TRN reporting given prospective registration

<!--# Are TRNs reported more or less often in abstract OR full text for prospective or retrospective registrations? -->

A chi-square test of independence was performed to examine the relation between prospective registration and TRN reporting in either the abstract or the full-text (abstract-only, full-text-only, both, neither). Registration was categorized as prospective if the trial was registered in the same or preceding months to trial start, since trial start on ClinicalTrials.gov is sometimes available with month and year only. Two trials (NCT02127736, NCT02213406) without start dates and were excluded from this analysis as we could not determine whether registration was prospective. The relation between these variables was significant, X2 (1, N = 1894) = 22.8, p = .00000182. Prospectively registered trials (Proportion reporting = X) were more likely than retrospectively registered trials (Proportion reporting = X) to report the TRN in either the abstract and/or the full-text (Difference in reporting = X, CI = X, X).

### Change in registration-publication linkage over time, by registry

Overall, DRKS trials appear to be less robustly linked than ClinicalTrials.gov trials. Overall, we observed positive trends for TRN reporting in full-text and abstracts. <!--# add test? --> Figure \@ref(fig:plot-links-year-reg) shows the registration-publication links over time for ClinicalTrials.gov and DRKS.

@@

>>>>

# Source analyses
source(here("R", "functions", "screening.R"))
# # source(here("R", "functions", "report-results.R"))
source(here("R", "analysis", "01_screen-trials.R"))
# # source(here("R", "analysis", "02_flowchart-trials.R"))
source(here("R", "analysis", "03_tbl-trials.R"))
source(here("R", "analysis", "04_link-reg-pub.R"))
# # source(here("R", "analysis", "05_upset-link-reg-pub.R"))
source(here("R", "analysis", "06_plot-change-over-time.R"))
source(here("R", "analysis", "trn-regex-analysis.R"))
source(here("R", "analysis", "cross-registrations.R"))
source(here("R", "analysis", "logistic-regression.R"))
# # fs::dir_ls(here("R", "analysis")) %>% 
# #   purrr::walk(source)
```

@@

**Secondary analyses** We constructed logistic regression models to explore the relationship between trial completion date, registry, and registration-publication linking. We also looked at reported cross-registrations in the registries as well as additional TRNs reported in trial publications.

We explored<!--# tested --> the relationships between (1) TRN in abstract or full-text, and TRN in PubMed metadata, as well as between (2) TRN in PubMed metadata and a linked publication in the registration. We also explored the relationship between registration-publication links and (1) registry, and (2) prospective registration <!--# not sure about this. i think if we do this then glm for the various practices -->. We looked at change in registration-publication linking practices over time. <!--# We also looked at reporting rates by UMCs, randomized controlled trials (RCTs), and ICMJE-status. We also explored the association of TRN reporting practices with other responsible trial research practices, namely prospective registration and summary results reporting. All analyses are exploratory. -->

@@

## Automated pipeline for data collection

@@

### Inclusion and exclusion criteria

@@

<!--# Using registry APIs, when available, and web scraping, we checked whether each unique TRN resolved to a valid trial registration. currently placeholder "TRUE" such that all TRNs resolve in the registries -->

<!--# Since i'm focusing iv trn only, should i change this section so not ALL the trns? then would have to change my backend work. or can keep this section but maybe not report all the details in text but just appendix? -->

## Analysis

We generated descriptive statistics on trial and publication characteristics, overall and by registry. We calculated the number and proportion of trials linked via the publication (full-text, abstract, metadata) and the registration. <!--#with each type of link and different combination of links.-->

To explore change over time, the relationships between types of links, and differences between registries, we ran exploratory logistic regressions for each link type, with all other link types as well as registry and year as explanatory variables. In particular, since PubMed metadata may be generated from the publication abstract or full-text, we examined the relationship between TRN reporting in either the abstract or full-text, and TRN inclusion in the metadata.

Additionally, since ClinicalTrials.gov registrations automatically reference publications with the TRN in the PubMed metadata (whereas DRKS does not), we examined the proportion of automatically vs. manually linked publications in ClinicalTrials.gov. To explore registry differences for only manually linked references, we excluded automated links and calculated the number and proportion of trials with the publication linked in the registration. <!--#add automatic links to regression for pub in registry? in which case would need to recode NA as FALSE. To explore whether automated links explains registry differences in publication links in the registry...-->

@@

The IntoValue trials dataset includes all trials led by a German UMC that were registered on ClinicalTrials.gov or DRKS and completed between 2009 and 2017 (n = `r n_trials_iv`). After applying our exclusion criteria, our sample included `r n_trials_ft_pdf` trials with `r n_pubs_unique` unique results publications indexed in PubMed and available as full-text, as some publications include results from more than one trial. Figure \@ref(fig:flow-trials) details the trial and publication screening. Table \@ref(tab:tbl-trials) shows summary descriptive details of the trials with results publications by registry. <!--# QUESTION: what summary info about trials and publications? I also include a long version of demographic data in case we want to include any of these, ideally with collapsed categories, in Table @ref(tab:tbl-trials-full) -->

@@

While per our inclusion criteria all trials in our sample (n = `r n_trials_ft_pdf`) were registered and had a publication, only `r report_n_pct_trials(n_trials_link_full)` trials had the most comprehensive registration-publication linking, meaning a linked publication in the registration as well as the TRN in the publication full-text, abstract, and PubMed metadata. Disregarding metadata, which is beyond a trialist's control, an additional `r n_trials_link_ft_abs_reg - n_trials_link_full` trials had comprehensive linking for a total of `r report_n_pct_trials(n_trials_link_ft_abs_reg)`. In contrast, we found `r report_n_pct_trials(n_trials_link_none)` trials with no links in either the registration or the publication. The most common linking practice was reporting of the TRN in the full-text only, accounting for `r report_n_pct_trials(n_trials_link_trn_ft_only)` trials. The inclusion of the various link types ranged from `r report_pct_num_denom_trials(n_trials_link_trn_abs)` in abstracts to `r report_pct_num_denom_trials(n_trials_link_trn_ft)` in full-text. Table \@ref(tab:tbl-link-by-reg) shows registration-publication links overall and by registry. <!--# point estimates in table should have CI, right? -->Figure \@ref(fig:plot-upset-links-reg-pub-registries) shows the intersection of registration-publication links across the various types for ClinicalTrial.gov and DRKS trials.<!--#shows the number of trials with each link type and combination of link types, overall and by registry.-->

@@

### Logistic regression

```{r tbl-trials-regression}
tbl_trials_regression %>% 
  gtsummary::as_flex_table() %>% 
  flextable::set_caption(caption = "Crude and adjuusted odds ratios for factors associated with publication-registration links")
```

Table \@ref(tab:tbl-trials-regression) shows the crude univariate and adjusted multivariate odds ratios (ORs) for each type of publication-registration link across all explanatory variables. Completion year did not have a strong relationship with linking practices, although trials completing more recently were significantly more likely report the TRN in the abstract (`r inline_text(cor_abs, variable = "completion_year", pattern = "cOR {estimate} [{ci}]")`; `r inline_text(aor_abs, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`) and full-text (`r inline_text(cor_ft, variable = "completion_year", pattern = "cOR {estimate} [{ci}]")`; `r inline_text(aor_ft, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`). Figure \@ref(fig:plot-links-year-reg) shows the rate of registration-publication links over time for ClinicalTrials.gov and DRKS.

With regards to the relationship between link types,<!--#in the adjusted analysis,--> trials with one type of link were generally more likely to have another type of link. <!--#However, trials with a TRN in the full-text were significantly less likely to have the publication linked in the registration (`r inline_text(aor_reg, variable = "has_iv_trn_ft_pdf", pattern = "aOR {estimate} [{ci}]")` ; `r inline_text(aor_ft, variable = "has_reg_pub_link", pattern = "aOR {estimate} [{ci}]")`).--> In particular, TRNs were significantly more likely to appear in the PubMed metadata if included in the abstract (`r inline_text(aor_si, variable = "has_iv_trn_abstract", pattern = "aOR {estimate} [{ci}]")`), but not significantly more likely if included in the full-text (`r inline_text(aor_si, variable = "has_iv_trn_ft_pdf", pattern = "aOR {estimate} [{ci}]")`). Futhermore, DRKS TRNs were significantly less likely than ClinicalTrials.gov TRNs to appear in the PubMed metadata (`r inline_text(aor_si, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`). Similarly, for trials with a TRN in either abstract or full-text, ClinicalTrials.gov TRNs appeared at a higher rate than DRKS TRNs in the metadata (62% vs. 18%). Table X shows the number and proportion of trials with a TRN in the metadata given TRN in the abstract, full-text, or either, overall and by registry.

**Table X. Number and Proportion of trials with TRN in metadata given TRN in abstract, full-text, or either. Denominator indicates number of trials with TRN in abstract or full-text, from the respective registry.**

|                                         | Overall        | ClinicalTrials.gov | DRKS         |
|-----------------------------------------|----------------|--------------------|--------------|
| **TRN in either abstract or full-text** | 53% (723/1372) | 62% (669/1077)     | 18% (54/295) |
| **TRN in abstract**                     | 85% (605/714)  | 97% (562/580)      | 32% (43/134) |
| **TRN in full-text**                    | 45% (513/1131) | 54% (471/878)      | 17% (42/253) |

DRKS trials were also significantly less likely than ClinicalTrials.gov trials to reference the publication in the registration (`r inline_text(aor_reg, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`). <!--#Inclusion of the TRN in the abstract was significantly less likely in DRKS than ClinicalTrials.gov trials in the crude data (`r inline_text(cor_abs, variable = "registry", level = "DRKS", pattern = "cOR {estimate} [{ci}]")`), whereas more likely in the adjusted analysis (`r inline_text(aor_abs, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`).--> Most (`r report_pct_num_denom_trials(n_trials_ctgov_link_reg_trn_si_derived, n_trials_link_reg_ctgov)`) ClinicalTrials.gov trials referencing the publication in the registration had the link automatically derived from the PubMed metadata, which is not currently possible for DRKS. Excluding trials with automatically linked publications, ClinicalTrials.gov and DRKS had similar rates of referencing publication in the registration (24%, 186/783 vs. 21%, 96/447).

@@

We found that `r report_pct_num_denom_trials(n_trials_link_reg)` of registrations included a reference with a publication identifier (DOI or PMID) to the associated publication in our sample of trials with published results. Only `r report_pct_num_denom_trials(n_trials_link_reg_drks, n_trials_drks)` of DRKS trials referenced the results publication in the registration, in contrast with `r report_pct_num_denom_trials(n_trials_link_reg_ctgov, n_trials_ctgov)` of ClinicalTrials.gov trials, of which `r report_pct_num_denom_trials(n_trials_ctgov_link_reg_trn_si_derived, n_trials_link_reg_ctgov)` were automatically indexed from PubMed metadata.

@@

We found that trials with a ClinicalTrials.gov TRN were significantly more likely than trials with a DRKS TRN to include a publication reference in the registry (`r apa::apa(chisq_link_by_reg_reg, print_n = TRUE)`) and to include the TRN in the PubMed metadata (`r apa::apa(chisq_link_by_reg_si, print_n = TRUE)`) and abstract (`r apa::apa(chisq_link_by_reg_abs, print_n = TRUE)`). We did not find a significant relationship between registry and TRN reporting in full-text (`r apa::apa(chisq_link_by_reg_pdf, print_n = TRUE)`).

@@

@@

# Sensitivity Analyses
